\appendix

\section{Protocol}

In this section, we describe the protocols that are implemented by our
library.

For each class, generic function, and condition type, we only include
a brief description.  In particular, we do not include a complete
description of the exceptional situations possible.  For a complete
description, see the \texttt{Documentation} subdirectory in
the repository at GitHub.%
\footnote{\texttt{https://github.com/robert-strandh/Cluffer}}

\subsection{Classes}

\Defprotoclass {buffer}

\vskip 1mm\noindent
This class is the base class of all buffers.  Each different buffer
implementation defines specific implementation classes to be
instantiated by client code.

\Defprotoclass {line}

\vskip 1mm\noindent
This class is the base class of all lines.  Each different line
implementation defines specific implementation classes to be
instantiated by client code.

\Defprotoclass {cursor}

\vskip 1mm\noindent
This class is the base class of all cursors.  Each different line
implementation defines specific implementation classes to be
instantiated by client code.

\subsection{Generic functions}

\Defgeneric {item-count} {entity}

\vskip 1mm\noindent
If \textit{entity} is a line, then return the number of items in that
line.  If \textit{entity} is a cursor, return the number of items in
the line in which \textit{cursor} is located.  If \textit{entity} is a
buffer, then return the number of items in the buffer.

\Defgeneric {item-at-position} {line position}

\vskip 1mm\noindent
Return the item located at \textit{position} in \textit{line}.

\Defgeneric {insert-item-at-position} {line item position}

\vskip 1mm\noindent
Insert \textit{item} into \textit{line} at \textit{position}.

After this operation completes, what happens to cursors located at
\textit{position} before the operation depends on the class of the
cursor and of \textit{line}.

\Defgeneric {delete-item-at-position} {line position}

\vskip 1mm\noindent
Delete the item at \textit{position} in \textit{line}.

\Defgeneric {cursor-position} {cursor}

\vskip 1mm\noindent
Return the position of \textit{cursor} in the line to which it is
attached.

\Defgeneric {(setf cursor-position)} {new-position cursor}

\vskip 1mm\noindent
Set the position of \textit{cursor} to \textit{new-position} in the
line to which \textit{cursor} is attached.

\Defgeneric {insert-item} {cursor item}

\vskip 1mm\noindent
Calling this function is equivalent to calling
\texttt{insert-item-at-position} with the line to which
\textit{cursor} is attached, \textit{item}, and the position of
\textit{cursor}.

\Defgeneric {delete-item} {cursor}

\vskip 1mm\noindent
Delete the item immediately after \emph{cursor}.

Calling this function is equivalent to calling
\texttt{delete-item-at-position} with the line to which
\textit{cursor} is attached and the position of \textit{cursor}.

\Defgeneric {erase-item} {cursor}

\vskip 1mm\noindent
Delete the item immediately before \emph{cursor}.

Calling this function is equivalent to calling
\texttt{delete-item-at-position} with the line to which
\textit{cursor} is attached and the position of \textit{cursor} minus
one.

\Defgeneric {cursor-attached-p} {cursor}

\vskip 1mm\noindent
Return \emph{true} if and only if \textit{cursor} is currently
attached to some line.

\Defgeneric {detach-cursor} {cursor}

\vskip 1mm\noindent
Detach \textit{cursor} from the line to which it is attached.

\Defgeneric {attach-cursor} {cursor line \optional (position 0)}

\vskip 1mm\noindent
Attach \textit{cursor} to \textit{line} at \textit{position}.

\Defgeneric {beginning-of-line-p} {cursor}

\vskip 1mm\noindent
Return \textit{true} if and only if \textit{cursor} is located at the
beginning of the line to which \textit{cursor} is attached.

\Defgeneric {end-of-line-p} {cursor}

\vskip 1mm\noindent
Return \textit{true} if and only if \textit{cursor} is located at the
end of the line to which \textit{cursor} is attached.

\Defgeneric {beginning-of-line} {cursor}

\vskip 1mm\noindent
Position \textit{cursor} at the very beginning of the line to which it
is attached.

\Defgeneric {end-of-line} {cursor}

\vskip 1mm\noindent
Position \textit{cursor} at the very end of the line to which it
is attached.

\Defgeneric {forward-item} {cursor}

\vskip 1mm\noindent
Move \textit{cursor} forward one position.

\Defgeneric {backward-item} {cursor}

\vskip 1mm\noindent
Move \textit{cursor} backward one position.
