\documentclass{sig-alternate-05-2015}
\usepackage[utf8]{inputenc}
\usepackage{color}

\def\inputfig#1{\input #1}
\def\inputtex#1{\input #1}
\def\inputal#1{\input #1}
\def\inputcode#1{\input #1}

\inputtex{logos.tex}
\inputtex{refmacros.tex}
\inputtex{other-macros.tex}

\begin{document}
\setcopyright{rightsretained}
\title{A CLOS Protocol for Editor Buffers}
\numberofauthors{1}
\author{\alignauthor
Robert Strandh\\
\affaddr{University of Bordeaux}\\
\affaddr{351, Cours de la Libération}\\
\affaddr{Talence, France}\\
\email{robert.strandh@u-bordeaux1.fr}}

\toappear{Permission to make digital or hard copies of all or part of
  this work for personal or classroom use is granted without fee
  provided that copies are not made or distributed for profit or
  commercial advantage and that copies bear this notice and the full
  citation on the first page. Copyrights for components of this work
  owned by others than the author(s) must be honored. Abstracting with
  credit is permitted. To copy otherwise, or republish, to post on
  servers or to redistribute to lists, requires prior specific
  permission and/or a fee.

%  ACM 978-1-4503-2931-6/14/08\$15.00.???
%  http://dx.doi.org/10.1145/2635648.2635656
}

\maketitle

\begin{abstract}
Many applications and libraries contain a data structure for storing
and editing text.  Frequently, this data structure is chosen in a
somewhat arbitrary way, without taking into account typical use cases
and their consequence to performance.  In this paper, we present a
data structure in the form of a \clos{} protocol (together with two
implementations of it) that addresses these issues.  In particular,
the protocol is divided into an \emph{edit} protocol and an
\emph{update} protocol, designed to be executed at different
frequencies.  The update protocol is based on the concept of
\emph{time stamps} allowing multiple \emph{views} without any need for
\emph{observers} or similar techniques for informing the views of
changes to the model (i.e., the text buffer).
\end{abstract}

\terms{Text editing}

\keywords{\clos{}, \commonlisp{}, Text editor}

\inputtex{spec-macros.tex}

\inputtex{sec-introduction.tex}
\inputtex{sec-previous.tex}
\inputtex{sec-our-method.tex}
\inputtex{sec-benefits.tex}
\inputtex{sec-conclusions.tex}
\inputtex{sec-acknowledgments.tex}
\inputtex{app-protocol.tex}

\bibliographystyle{abbrv}
\bibliography{cluffer}
\end{document}

%%  LocalWords:  sandboxing runtime
